---
title: "Day 8 EOD Activity"
format: html
---
# 1. Setup 
```{r}
library(here)
library(tidyverse)
library(janitor)
library(scales)
```


```{r}
# read in data 
space <- read_csv(here("data", "space_launches.csv"))

# looking at data 
# numeric = launch_year 
# character = type, mission, agency, 
```

# Task 1: Do your data viz worst
```{r}
ggplot(space) +
  geom_hex(aes(x = launch_year, y = agency)) +
  scale_fill_gradient2() +
  labs(x = "launch year",
       y = "where it is from",
       title = "rockets") +
  theme(panel.grid.major = element_line(color = "#a67289"),
        plot.background = element_rect(fill = "#ffe700"),
        text = element_text(size = 10, family = "Comic Sans MS"), 
        axis.text.x = element_text(color = "pink"),
        axis.text.y = element_text(color = "purple"),
        panel.background = element_rect(color = "#0044b8", fill = "#3e4616"),
        axis.text.y.left = element_text(angle = 50))

ggsave(here("figs", "jordan_badplot.png"))
       
```


# Task 2: Wrangling practice with the SF Greenhouse Gas data

```{r}
# read in data 
sf_greenhouse <- read_csv(here("data", "San_Francisco_Communitywide_Greenhouse_Gas_Inventory.csv"))

sf_clean <- janitor::clean_names(sf_greenhouse)

```


1. Create a summary table of total annual greenhouse gas emissions (only from Emissions_mtCO2e column) by sector_general
```{r}
task2.1 <- sf_clean %>% 
  group_by(sector_general) %>% 
  summarise(total_emissions = sum(emissions_mt_co2e, na.rm = TRUE))
  
```

2. Create a summary table of total annual greenhouse gas emissions (only from Emissions_mtCO2e column) by year and commodity type
```{r}
task2.2 <- sf_clean %>% 
  group_by(calendar_year, commodity_type) %>% 
    summarise(total_emissions = sum(emissions_mt_co2e, na.rm = TRUE))

```

3. Create a subset of the data that only contains observations from 2005 on, only for observations where Sector_Detail2 contains the pattern “PG&E”, then limit to columns Calendar_Year, Sector_Detail2 and Emissions_mtCO2e.

```{r}
# year >= 2005 
# sector_detail2 contains "PG&E" think str_detect
# only calendar_year, sector_detail2, emissions_mtco2e

task2.3 <- sf_clean %>% 
  filter(calendar_year >= 2005) %>% 
  filter(str_detect(sector_detail2, "PG&E")) %>% 
  select(calendar_year, sector_detail2, emissions_mt_co2e)
```


# Task 3: Do your data viz best

You get to choose what you want to visualize. You can pick which variables & observations you’re interested in. You may want to do some wrangling / summarizing first.

- I want to do total emissions by sector in the year 2019 
- bar graph sector_general by total_emissions (i need to make this)

```{r}
graph_subset <- sf_clean %>% 
  filter(calendar_year == 2019) %>% 
  group_by(sector_general) %>% 
  summarise(total_emissions = sum(emissions_mt_co2e, na.rm = TRUE)) 
```

```{r}
# plot it 

ggplot(data = graph_subset, aes(y = sector_general, x = total_emissions, 
                                fill = sector_general,)) +
  geom_col(fill = "darkolivegreen", show.legend = FALSE) +
  labs(x = "Total carbon emissions\n(mtCO2e)\n",
       y = "Sector",
       title = "San Francisco greenhouse gas emissions (2019)",
       caption = "Data: San Francisco Communitywide Greenhouse Gas Inventory") +
  scale_x_continuous(labels = comma) +
  theme_linedraw()
  
ggsave(here("figs", "jordan_sf_plot.png"))
```


